Named Entity Recognition using Structured Prediction

### 2012-11-06 doc/report/paper.tex
Report++.

@@

\begin{table}[h!]
\scriptsize
\begin{tabular}{l|l}
\hline
\bf Named Entity Type & \bf Example \\
\hline
Person Names & Greta Garbo, John, Mary \\
Organisations& Benetton, Coca-Cola Gmbh\\
Locations&  Paris, Australia, Bristol\\
Miscellaneous& World Cup\\
 Date Expression& 01-02-2012, January, 6th, 1998 \\
Time Expression & 10 p.m.\\
Monetary Value &  \$15, 100 Euro   \\
Percent &   30\%
\end{tabular}
\caption{NE Examples}
\label{table:NETypes}
\end{table}

@@

\subsection*{CoNLL Data Set \& Baseline Performance}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.

The baseline for \emph{Named Entity Recognition} in the four different languages is visualised in table \
\begin{table}[h!]
\scriptsize
\begin{tabular}{l|l|l|l}
\hline
\bf Language & \bf Precision & \bf Recall & \bf $F_1$-measure \\ \hline
Spanish &             26.27 \% & 56.48 \% & 35.86  \%        \\
Dutch  &             64.38 \%  &45.19   \%    & 53.10 \%  \\
English &              71.91\%& 50.90 \%  & 59.61 $\pm$ 1.2 \%\\
German &      31.86 \%  & 28.89 \% & 30.30  $\pm$ \%1.3 \\
\end{tabular}
\caption{NER Baseline}
\label{table:Base}
\end{table}

@@

\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} 0  0  ORG  0  0 0 0 0 PER PER.  \label{PredEx1b}

@@

\section{Our Implementation}
% problem external data / use Dbpedia - gazetteer
Our NER system for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
respectively.

\paragraph*{Learning}
Labels of the whole sentence (0) for zero entity

\paragraph*{Structure}
Structured Perceptron with Averaging

\paragraph*{Decoding}
In this system we are using viterbi decoding, thus we are looking for the prediction $\hat{y}$, which is the sequence with the highest overall score:

\[
\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
\]

This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ using the Viterbi algorithm. 
As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
The score of the best sequence ending in $a$ is:

@@

\noindent And $\delta_i(a)$ can be calculated recursively:

@@

\subsection*{Features}
The features employed in the system can be divided into three categories: node, label and gazetteer features. 
We describe each of the three groups in the following.

>>>>

\newcommand{\namedentity}{Named Entity} 
\newcommand{\Oo}{\texttt O}

@@

\begin{table}[h!]
\scriptsize
\begin{tabular}{l l}
\bf Named Entity Type & \bf Example \\
\hline
Person Names & Greta Garbo, John, Mary \\
Organisations& Benetton, Coca-Cola Gmbh\\
Locations&  Paris, Australia, Bristol\\
Miscellaneous& World Cup\\
 Date Expression& 01-02-2012, January, 6th, 1998 \\
Time Expression & 10 p.m.\\
Monetary Value &  \$15, 100 Euro   \\
Percent &   30\% \\
\end{tabular}
\caption{NE Examples}
\label{table:NETypes}
\end{table}

@@

\subsection*{CoNLL Data and Baseline Performance}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.

The baseline for \emph{Named Entity Recognition} in the four different languages is visualised in table \
\begin{table}[h!]
\scriptsize
\begin{tabular}{|l|l|l|l|}
\hline
\bf Language & \bf Precision & \bf Recall & \bf $F_1$-measure \\ \hline
Spanish &             26.27\% & 56.48\% & 35.86\%        \\
Dutch  &             64.38\%  &45.19\%    & 53.10\%  \\
English &              71.91\%& 50.90\%  & 59.61 $\pm$ 1.2\%\\
German &      31.86\%  & 28.89\% & 30.30  $\pm$ 1.3\% \\
\hline
\end{tabular}
\caption{NER Baseline}
\label{table:Base}
\end{table}

@@

\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} O  O  ORG  O  O O O O PER PER.  \label{PredEx1b}

@@

\section{Implementation}

\subsection{Learning and Decoding}

% problem external data / use Dbpedia - gazetteer
Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets respectively.

As the predicted structure, we use the labels for every token in the full sentence. If a token is not a \namedentity, it is assigned the label \Oo.

Learning is implemented with a Structured Perceptron. To avoid overfitting, we average the parameters after the last iteration. For finding the best sequence of labels, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:

\[
\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \sum_{i=1}^{n}\mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
\]

This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ via dynamic programming. 
As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
The score of the best sequence ending in $a$ is:

@@

\noindent This function $\delta_i(a)$ can be defined recursively as:

@@

\subsection{Features}
The features employed in the system can be divided into three categories: node, label and gazetteer features. 
We describe each of the three groups in the following.
